---
title: "Data Science in Statistical Methods using R"
author: "Md Sayeef Alam"
date: "21/09/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Day 1

## Session 1: Application of Regression and Multiple Regression in Data Science

### Dr. R. K. Jana, IIM Raipur

Simple addition in R

```{r c001}
1+1
```

Some packages to be installed

```{r c002, results='hide', error = F, message=F, warning=F}
install.packages("matlib", dependencies = T,repos = "http://cran.us.r-project.org")
install.packages("corpcor", dependencies = T,repos = "http://cran.us.r-project.org") 
install.packages("GPArotation", dependencies = T,repos = "http://cran.us.r-project.org")
install.packages("psych", dependencies = T,repos = "http://cran.us.r-project.org")
install.packages("FactoMineR", dependencies = T,repos = "http://cran.us.r-project.org")
install.packages("tseries", dependencies = T,repos = "http://cran.us.r-project.org")
install.packages("corrplot", dependencies = T,repos = "http://cran.us.r-project.org")
install.packages("tseries", dependencies = T,repos = "http://cran.us.r-project.org")
install.packages("ggpubr", dependencies = T,repos = "http://cran.us.r-project.org")
install.packages("tidyverse", dependencies = T,repos = "http://cran.us.r-project.org")
install.packages("Hmisc", dependencies = T,repos = "http://cran.us.r-project.org")
install.packages("dplyr", dependencies = T,repos = "http://cran.us.r-project.org")
install.packages("ggplot2", dependencies = T,repos = "http://cran.us.r-project.org")
install.packages("lattice", dependencies = T,repos = "http://cran.us.r-project.org") 
install.packages("grid", dependencies = T,repos = "http://cran.us.r-project.org")
install.packages("DMwR", dependencies = T,repos = "http://cran.us.r-project.org")
install.packages("stats", dependencies = T,repos = "http://cran.us.r-project.org")
install.packages("nortest", dependencies = T,repos = "http://cran.us.r-project.org")
```

Adding the libraries corresponding to packages.

```{r c003,results='hide', error = F, message=F}
library(dplyr)
library(tseries)
library(matlib)
library(corpcor) 
library(GPArotation) 
library(psych)
library(FactoMineR)
library(corrplot)
library(ggpubr)
library(lattice)
library(grid)
library(nortest)
library(stats)
library(DMwR)
library(ggplot2)
```

Reading xls and xlsx files

```{r c004, eval = F}
install.packages("gdata", dep = T,repos = "http://cran.us.r-project.org")
library(gdata)
xls.data = read.xls("file.xls")
```

You need to specify the sheetIndex (sheet number) 

```{r c005, eval = F}
install.packages("xlsx", dep = T,repos = "http://cran.us.r-project.org") 
library(xlsx)
xlsx.data = read.xlsx("file.xlsx", sheetIndex = 1)
```

## Linear Regression
Simple Linear Regression

1 dependent (y)

1 independent (x)

Assumptions

1. Relationships between the above two must be linear

2. Residuals should be normally distributed

3. Residuals should be homoscedastic

4. Residuals should be independent

Homoscedasticity means same variance, error term (i.e. distance of the points from the fitted line) should be same across all values of the independent variables. 

Heteroscedasticity is when the error varies with the values of the independent variables.

Several measures are there to check for homoscedasticity

```{r c006}
library(datasets)
data(cars)
```

Lets check the variables inside the dataset

```{r c007}
names(cars)
head(cars)
```

Lets plot some parameters specifically speed vs distance

```{r c008}
plot(cars$speed, cars$dist)
plot(cars$speed, cars$dist, xlab = "Car Speed at Stopping (in mph)", 
     ylab = "Stopping Distance (in feet)", main = "The Effect of Car Speed on Stopping Distance")
```

Fitting a smooth line

```{r c009}
scatter.smooth(x=cars$speed, y=cars$dist, main="Dist ~ Speed")
```

Density plots for speed and distance

```{r c010}
library(e1071)
par(mfrow=c(1, 2))

plot(density(cars$speed), main="Density Plot: Speed", ylab="Frequency", sub=paste("Skewness:", round(e1071::skewness(cars$speed), 2)))  
polygon(density(cars$speed), col="red")

plot(density(cars$dist), main="Density Plot: Distance", ylab="Frequency", sub=paste("Skewness:", round(e1071::skewness(cars$dist), 2)))
```

Linear regression model fitting

```{r c011}
carmod <- lm(dist ~ speed, data = cars)
summary(carmod)
```

95% CI

```{r c012}
confint(carmod, level = 0.95)
```

Normality of residuals check

```{r c013}
res = carmod$residuals
hist(res)
```

### Interpretation
The coefficients in linear regression model states that with a unit change in x how much change is expected in y.

# Session 2: Data Science & Sample Survey

## Prof. G. N. Singh, IIT (ISM) Dhanbad

Word Statistics

In a literal sense 

Plural sense some sort of data numerical figures in our day to day arising, runs and all figures are called statistics

In singular collection of methods and principles in a book, 

Procedure to collection, analyse and interpret the data is called statistics

Statistics never claims 100% accuracy

Statistics is the science of decision making. As no decision is free from error.

Hope that PPTs will be provided soon.

# Day 2

## Session 3: Introduction to R

### Prof. G. N. Singh, IIT (ISM) Dhanbad

Theory and PPT will be available.

## Session 4: Introduction to R

### Dr. Anup Kumar Sharma, NIT Raipur

Theory and PPT will be available.

## Session 5: Graphical representation and normality testing in R

### Dr. Dhaval Maheta, 

```{r c014}
mtcars
attach(mtcars)
```

Find the mean for all columns

the number in between is the parameter denoting the 1 = row and 2 = column. row mean is useless so we are looking at column mean

```{r c015}
apply(mtcars,2,mean)
```

Now similarly for median and mode

```{r c016}
apply(mtcars,2,median)
apply(mtcars,2,mode)
```

aggregate function helps to calculate the required function (mean/median/mode/sd) for each category of independent variable

```{r c017}
aggregate(mpg~am,FUN = mean)
aggregate(mpg~am,FUN = median)
aggregate(mpg~am,FUN = mode)
aggregate(mpg~am,FUN = sd)
```

Find 3 way table to summary statistics and describeBy (available in psych library)

```{r c018}
aggregate(mpg~am+vs,FUN = mean)
summary(mtcars)
describeBy(mpg,am)
```

best descriptive summarizer called the stargazer, the flip = T command helps to transpose the rows and columns

```{r c019, error = F, message=F, warning=F}
install.packages("stargazer", dependencies = T,repos = "http://cran.us.r-project.org")
library(stargazer)

stargazer(mtcars,type = "text", title = "Descriptive Stats", digits = 1)
stargazer(mtcars,type = "text", title = "Descriptive Stats", digits = 1, flip = T)
```

Try the following codes to obtain data like SPSS

```{r c020, error = F, message=F, warning=F}
install.packages("summarytools", dependencies = T,repos = "http://cran.us.r-project.org")
install.packages("ellipsis", dependencies = T,repos = "http://cran.us.r-project.org")
library(summarytools)
library(ellipsis)
```

```{r c021}
attach(mtcars)
summarytools::descr(mtcars)
summarytools::freq(am)
summarytools::ctable(am,vs)
summarytools::dfSummary(mtcars)
```

Graphical representation of data

Using a new data set called "Orange"

```{r c022}
Orange
attach(Orange)
```

Density plot of circumference

```{r c023}
densityplot(~circumference)
```

Converting into categorical/factor variable

```{r c024}
tree.f = factor(Tree, levels = c(1,2,3,4,5), labels = c("tree1","tree2","tree3","tree4","tree5"))
```

Boxplot of age of trees

```{r c025}
bwplot(~age|tree.f)
```

Dot plot

```{r c026}
dotplot(age~circumference|tree.f)
```

Level plot

```{r c027}
levelplot(tree.f~circumference*age)
```

```{r c028, error = F, message=F, warning=F}
install.packages("DataExplorer", dependencies = T,repos = "http://cran.us.r-project.org")
library(DataExplorer)

plot_histogram(mtcars)
```

Not even interested to write a single line of command, this is very sexy and appealing for data cleaning

```{r c029, error = F, message=F, warning=F}
install.packages("esquisse", dependencies = T,repos = "http://cran.us.r-project.org")
library(esquisse)

esquisse::esquisser(mtcars)
```

Even new packages click and play, contingency tables, summary stats

```{r c030, error = F, message=F, warning=F}
install.packages("Rcmdr", dependencies = T,repos = "http://cran.us.r-project.org")
library(Rcmdr)
```

Lets use our own dataset Employee dataset but for now use mtcars

```{r c031}
mtcars
attach(mtcars)
```

Normality Checks using graphics but graphics is not 100% so we use test for rejection of h0: normal distribution and h1: not normal ; hence if p value < 0.05 then the data is not normal

```{r c032}
qqnorm(mpg)
qqline(mpg)
shapiro.test(mpg)
```

tm, quanteda: for unstructured data

tseries: for timeseries

animate: can be used to animate any plot type, written by Yihui Xie

gganimate: used to specifically animate ggplot graphics, written by Thomas Lin Pedersen

plotly: an interactive plotting library which has animation features

googlevis: has a flash based motion chart option

plspm for SEM
